---
title: "Lecture Assignments - Using Models 1"
author: "Jordan Smith"
date: "2022-11-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# {.tabset}

## Prep Work
```{r}
require("here")
catrate = read.csv(here("data", "catrate.csv"))
head(catrate)
summary(catrate)
```

**Histogram of Catastrophe Rates**

```{r}
hist(catrate$cat.rate,
     main = "Histogram of Catastrophe Rates",
     ylab = "Catastrophe Rate",
     xlab = "Frequency")
```

**Checking for Normality**

```{r}
#Shapiro Test
shapiro.test(catrate$cat.rate)

#Other Normality Tests
install.packages("nortest", repos = "http://cran.us.r-project.org")
library(nortest)

ad.test(catrate$cat.rate)
lillie.test(catrate$cat.rate)
```

**One-Sample Tests**

```{r}
#Parametric One-Sample Test: The t-test
t.test(x = catrate$cat.rate, mu = 0.2857143)

#One-Sided Alternative Hypothesis
t.test(x = catrate$cat.rate, mu = 0.28571413, alternative = "greater")

#Non-Parametric One-Sample Test: The Wilcoxon Rank Sum Test
wilcox.test(catrate$cat.rate, mu = 2/7, exact = FALSE)

#One-Sided Alternative Hypothesis with The Wilcoxon Rank Sum Test
wilcox.test(catrate$cat.rate, mu = 2/7, alternative = "greater")
```

**Two-Sample Tests (Using Palmer Penguins)**

```{r}
require(palmerpenguins)
penguin_dat = droplevels(subset(penguins, species !="Gentoo"))
summary(penguin_dat)

#Visualizing a Conditional Boxplot
boxplot(
  flipper_length_mm ~ species,
  data = penguin_dat,
  ylab = "Flipper Length (mm)")

#Testing for Normality
##Has to be done for each individual species
dat_adelie = subset(penguin_dat, species == "Adelie")
shapiro.test(dat_adelie$flipper_length_mm)

wilcox.test(flipper_length_mm ~ species,
            data = penguin_dat)

dat_chinstrap = subset(penguin_dat, species == "Chinstrap")
shapiro.test(dat_chinstrap$flipper_length_mm)

#Parametric and NonParametric Tests
t.test(flipper_length_mm ~ species,
       data = penguin_dat)

```

## Question 1
**1. Create a histogram of the salamander reproduction catastrophic rates.**
```{r}
hist(catrate$cat.rate,
     main = "Histogram of Catastrophe Rates",
     ylab = "Catastrophe Rate",
     xlab = "Frequency")
```

## Question 2 - 4

**2. Conduct a Shapiro-Wilk test of normality of the salamander catastrophic rates. Report the p-value and show the R-code you used to conduct the test.**
```{r}
shapiro.test(catrate$cat.rate)
```

**3. What is the null hypothesis for the Shapiro test?**
The null hypothesis states that the given data was sampled from a normally-distributed population.

**4. Based on the Shapiro test results, is there strong evidence that the sample came from a non-normally distributed population?**
With a p-value of 0.04097, there is statistically significant evidence that the null-hypothesis can be rejected and the sample did not come from a normally distributed population.

## Question 5 - 7
*Conduct a one-sample t-test of the alternative hypothesis that the catastrophic rate is different from the pond-filling rate.*

**5. Show the code you used to conduct the t-test.**
```{r}
t.test(x = catrate$cat.rate, mu = 0.2857143)
```

**6. State the null hypothesis of the test, in plain nontechnical English.**
The null hypothesis states that there is no difference in the catastrophic rate and the late-filling rate of marbled salamanders.

**7. Is this a one- or two-tailed test?**
This is a two-tailed test since we are just checking to see if the catastrophic rate is different, not specifically greater or less than, the late-filling rate.

## Question 8 - 10

**8. What is the p-value from your t-test? Interpret the p-value as a false-positive rate using nontechnical English.**
The p-value from this t-test is 0.01193, which from a false-positive perspective tells us that we would have incorrectly rejected the hypothesis that the two averages were different when in reality there was no statistical difference between them.

**9. What is the confidence interval for the difference between the null hypothesis and alternative hypothesis means? Did it include zero?**
The confidence interval between the two is 0.3526250 - 0.7261295 and it does not include zero.

**10. Considering the results from your t-test, did you conclude that there was strong evidence to reject the null hypothesis?**
Looking at the p-value of 0.01193, it is below p = 0.05 which tells us that we have grounds to reject the null hypothesis.

## Question 11 - 13
*Conduct a one-sample Wilcoxon rank sum test of the alternative hypothesis that the catastrophic rate is different from the pond late-filling rate. Interpret the results of your one-sample Wilcoxon rank sum test of the alternative hypothesis that the catastrophic rate is different from the pond-filling rate.*

**11. Show the code you used to conduct the test**
```{r}
wilcox.test(catrate$cat.rate, mu = 2/7, exact = FALSE)
```

**12. Compare the p-value with the p-value you got from the t-test.**

t-test p-value = 0.01193
Wilcoxon Test p-value = 0.006275

**13. Considering the results from your rank sum test, did you conclude that there was strong evidence to reject the null hypothesis?**
Since the p-value from the rank sum test was below p = 0.05, we have evidence to reject the null hypothesis that the two means are the same.

## Question 14 - 15
*Consider the one-sample t-test and Wilcoxon rank sum tests that you performed on the catastrophic rate data. Explain your reasoning*

**14. Compare the overall conclusions you could draw from the results of the two tests.**
These two tests give us significant evidence that the means of the catastrophic rate and the late-filling rate are different, and that there may be additional factors in place to justify the catastropic rate of marbled salamanders.

**15. Considering the numerical and graphical data exploration, which test do you think was more appropriate for these data?**
I think the t-test and wilcoxon rank sum tests were the most appropriate because they had outputs of specific values rather than an overal visual display.

## Question 16 - 17

**16. Show the R-code you used to conduct tests of normality for the flipper lengths of Chinstrap and Adelie penguins.**
```{r}
require(palmerpenguins)
penguin_dat = droplevels(subset(penguins, species !="Gentoo"))

dat_adelie = subset(penguin_dat, species == "Adelie")
shapiro.test(dat_adelie$flipper_length_mm)

dat_chinstrap = subset(penguin_dat, species == "Chinstrap")
shapiro.test(dat_chinstrap$flipper_length_mm)

wilcox.test(flipper_length_mm ~ species,
            data = penguin_dat)

```

**17. Interpret the test results. Do you conclude that the flipper lengths are normally-distributed for each species? Make sure your answers make reference to the test output.**
Looking at the results of the Shapiro test for each of the species, Adelie penguins had a p-value = 0.72 and Chinstrap had a p-value = 0.81, both of these values indicating that the distribution of flipper lengths is normally distributed for both species.

## Question 18
*Create a single figure consisting of histograms of flipper lengths of Adelie and Chinstrap penguins. Your figure should have two histograms arranged in one row.*

**18. Save your figure to a file and include it to your report. Your figure needs to have appropriate dimensions such that the two histograms are not vertically stretched**
```{r}
par(mfrow = c(1,2))
hist(dat_adelie$flipper_length_mm,
     main = "Adelie Flipper Length",
     xlab = "Flipper Length",
     ylab = "Frequency")
hist(dat_chinstrap$flipper_length_mm,
     main = "Chinstrap Flipper Length",
     xlab = "Flipper Length",
     ylab = "Frequency")
```

## Question 19 - 20
*Conduct a two-sample t-test of the alternative hypothesis that the Adelie penguins have different flipper lengths than chinstrap penguins.*

**19. State the alternative hypothesis of the test, in plain nontechnical English**
The alternative hypothesis for this test is that the flipper lengths of Adelie and Chinstrap penguins is different.

**20. Include the code you used to conduct the t-test.**
```{r}
t.test(flipper_length_mm ~ species, data = penguin_dat)
```