---
title: "Probability and Frequentist Concepts"
author: "Jordan Smith"
date: "2022-10-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# {.tabset}

## Question 1
What is the probability of observing a count of exactly 3 successes in a binomial distribution with parameters n = 4 and p = 0.75?
``` {r}
dbinom(3, size = 4, prob = 0.75)
```
The probability of exactly 3 successes is 0.421875

## Question 2
What is the probability of observing a count of 3 successes or fewer in a binomial distribution with parameters n = 4 and p = 0.75?
```{r}
pbinom(3, size = 4, prob = 0.75)
```
The probability of 3 successes or fewer is 0.6835937

## Question 3
What is the probability of observing more than 3 successes in a binomial distribution with parameters n = 5 and p = 0.75?
``` {r}
1 - pbinom(3, size = 4, prob = 0.75)
```
The probability of more than 3 successes is 0.3164063

## Question 4
What is the probability of observing a value less than 1.2 from a normally distributed population with mean = 2 and standard deviation = 2?
```{r}
pnorm(1.2, mean = 2, sd = 2)
```
The probability of observing a value less than 1.2 is 0.3445783

## Question 5
What is the probability of observing a value of greater than 1.2 from a normally-distributed population with mean = 2 and standard deviation = 2?
```{r}
1 - pnorm(1.2, mean = 2, sd = 2)
```
The probability of observing a value greater than 1.2 is 0.6554217

## Question 6
What is the probability of observing a value between 1.2 and 3.2 from a normally-distributed population with the mean = 2 and standard deviation = 2?
```{r}
pnorm(3.2, mean = 2, sd = 2) - pnorm(1.2, mean = 2, sd = 2)
```
The probability of observing a value between 1.2 and 3.2 is 0.3811686

## Question 7
As the sample button is repeatedly hit, the histogram begins to follow the shape of the original distribution.

## Question 8
As the sample button is repeatedly hit, the histogram follows the trend of the original distribution, but seems to have a smaller range than the histogram using a sample size of 1.

## Question 9
As the sample button is repeatedly hit, the histogram continues to increase more, closer to the mean of the original distribution, and again having a smaller range compared to the histogram using.

## Question 10
There is such a drastic change between sample sizes 1 and 2 because you are essentially doubling the sample size, causing a huge shift in the accuracy of the data.

## Question 11
The two main factors are the sample size and confidence level.

## Question 12
How many 3-character words are possible?
```{r}
25^3
```
15625 3-character words are possible

## Question 13
B = 25^((410 * 40 * 80) + 1)
B = 25^1,312,001
