---
title: "Week 6 Questions"
author: "Jordan Smith"
date: "2022-10-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# {.tabset}

## Question 1
In the baseline scenario of seed predation, if there were no relationship between the seeds being removed across different species, there would be an equal predation rate when comparing the various species. Using the example of seed predation, a null hypothesis can be determined as no relationship between the predation rates of the two species of seeds. 

## Question 2

``` {r}
rm(list = ls())

pol_n_predation = 26
pol_n_no_predation = 184
pol_n_total = 210
pol_predation_rate = 0.124

psd_n_predation = 25
psd_n_no_predation = 706
psd_n_total = 731
psd_predation_rate = 0.034

print(
  paste0(
    "The seed predation rate for Polyscias fulva is: ",
    round(pol_predation_rate, digits = 3))) 

print(
  paste0(
    "The seed predation rate for Pseudospondias microcarpa is: ",
    round(psd_predation_rate, digits = 3)))
```

## Question 3
```{r}
#pol and psd vectors
pol = c(26, 184, 210, 0.124)
psd = c(25, 706, 731, 0.034)
pol_psd = c(pol, psd)

#table creation
seed_pred_table = matrix(pol_psd, byrow = TRUE, nrow = 2)

#table labels
seed_row_names = c("Polyscias fulva (pol)", "Pseudospondias microcarpa (psd)")
seed_col_names = c("Any Taken", "None Taken", "N", "Predation Rate")
rownames(seed_pred_table) = seed_row_names
colnames(seed_pred_table) = seed_col_names

#final table
seed_pred_table
```

## Question 4

``` {r}
#Ratio of Seed Predation Proportions = proportion for pol/proportion for psd
0.124/0.034
```
